---
title: "week-8-hw-eeb177"
author: "Krishna Krishnamurthy"
date: "March 4, 2017"
output: md_document
---

## Canids continued; Mapping stratigraphic records

```{r}
library("ggplot2")
canids <- read.csv("/home/eeb177-student/Desktop/eeb177/homework/sorted-canid-ranges.csv", header = F, as.is = T)
names(canids) <- c("genus", "species", "minage", "maxage")
head(canids)

canid_occ <- ggplot(canids, aes( species, ymin = maxage, ymax=minage, colour = genus))# plot data
canid_occ <- canid_occ + geom_linerange() # draw ranges
canid_occ <- canid_occ + theme(legend.position="none") # remove legend
canid_occ <- canid_occ + coord_flip() # flip labels
canid_occ <- canid_occ +  theme(axis.text.y = element_text(size=3)) # make species names smaller
canid_occ <- canid_occ + theme(axis.ticks.y=element_blank()) # remove tick marks
canid_occ <- canid_occ + scale_y_continuous(limits=c(0, 40), expand = c(0, 0), breaks=c(0, 10, 20, 30, 40)) # reduce empty space
canid_occ <- canid_occ + labs(title = "Canid Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20)) #adding title and modifying axes
canid_occ
ggsave(filename = "canid-occ.pdf", plot = canid_occ) # saving file in pdf format
```

## A report on canid fossil records
The figure shows the fossil records for canids. There are 211 unique species records and 206 unique genera. *Cynodictis lacustris* has the longest fossil range (35.55 Ma)

```{r}
library("ggplot2")
felines <- read.csv("/home/eeb177-student/Desktop/eeb177/homework/feline-ranges.csv", header = F, as.is = T)
names(felines) <- c("genus", "species", "minage", "maxage")
head(felines)

feline_occ <- ggplot(felines, aes( species, ymin = maxage, ymax=minage, colour = genus))# plot data
feline_occ <- canid_occ + geom_linerange() # draw ranges
feline_occ <- canid_occ + theme(legend.position="none") # remove legend
feline_occ <- canid_occ + coord_flip() # flip labels
feline_occ <- canid_occ +  theme(axis.text.y = element_text(size=3)) # make species names smaller
feline_occ <- canid_occ + theme(axis.ticks.y=element_blank()) # remove tick marks
feline_occ <- canid_occ + scale_y_continuous(limits=c(0, 40), expand = c(0, 0), breaks=c(0, 10, 20, 30, 40)) # reduce empty space
feline_occ <- canid_occ + labs(title = "Feline Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20)) #adding title and modifying axes
feline_occ
ggsave(filename = "feline-occ.pdf", plot = feline_occ) # saving file in pdf format
```

## A report on feline fossil records
The figure shows the fossil records for felines. There are 139 unique species records and 34 unique genera. Cynodictis lacustris has the longest fossil range (35.55 Ma). *Miracinonyx inexpectatus* has the longest fossil range (13.28 Ma)

```{r}
library("paleotree")
data(graptPBDB)
head(graptOccPBDB)
#get formal genera
occSpecies<-taxonSortPBDBocc(graptOccPBDB, rank="species")
#plot it!
plotOccData(occSpecies)
occGenus<-taxonSortPBDBocc(graptOccPBDB, rank="genus")
#plot it!
plotOccData(occGenus)
#taxicDivDisc with the retiolinae dataset
data(retiolitinae)
taxicDivDisc(retioRanges)
```

## Exercise 1: Exploring Data Frames

```{r}
# Adding rows and colums
# First we read in the data file we created for the class challenge

cats <- read.csv("/home/eeb177-student/Desktop/eeb177/class-assignments/data/feline-data.csv")
cats

# We want to add ages
age <- c(4,5,8)
cats <- cbind(cats,age)
cats

# Adding a new row
newRow <-list("tortoiseshell", 3.3, TRUE, 9)
cats <- rbind(cats, newRow)
# produces a NA value

levels(cats$coat) <- c(levels(cats$coat), "tortoiseshell")
# adding a new factor, otherwise R won't add the new row
cats <- rbind(cats, list("tortoiseshell", 3.3, TRUE, 9))
cats

# Removing rows
cats[-4,]
# we can also use the na.omit function
cats <- na.omit(cats)

# Appending data frame
cats <- rbind(cats, cats)
cats
# generates complicated row names

rownames(cats) <- NULL
cats
# simplifies the row names
```

```{r}
# Challenge 1
df <- data.frame(first = c('Krishna'), last = ("Krishnamurthy"), lucky_number = c(0), stringsAsFactors = FALSE)
df <- rbind(df, list("Irina", "Bird", 14))
df <- cbind(df, coffeetime = c(FALSE, TRUE))
df

# Challenge 2
source(file = "scripts/load-gapminder.R")

#script is:
download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv", destfile = "data/gapminder-FiveYearData.csv")
gapminder <- read.csv(file = "data/gapminder-FiveYearData.csv")

# Challenge 3
str(gapminder)
# gapminder is a data frame with 6 columns, and 1704 observations. County and continent are factors; year is an integer value, pop, lifeExp and gdpPercap are all numeric vectors
```


## Exercise 2: Subsetting Data

```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
x

# accessing indices
x[1]
x[4]

# multiple indices at once and multiple times
x[c(1,3)]
x[c(1,1,3,1)]

# slices
x[1:4]

x[6] # does not exist in this list; hence we get a NA
x[0] # never exists

# returning all values except one
x[-2]
x[-(1:3)] # exclude elements 1 through 3
x <- x[-4]
x
```

```{r}
# Challenge 1
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

x[2:4]
x[-c(1,5)]
x[c(2,3,4)]

# Challenge 2
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
x[-which(names(x) == "g")]
# A and C are correct;names(x) == "g" does not return any true values because there were none passed to the command. 

# Challenge 3
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
x_subset <- x[x<7 & x>4]
print(x_subset)

# Challenge 4
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
m[2,c(4,5)] # this extracts 11 and 14

# Challenge 5
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
# we can extract 2 from xlist with the following commands
xlist$b[2]
xlist[[2]][2]
xlist[["b"]][2]

# Challenge 6
mod <- aov(pop ~ lifeExp, data=gapminder)
attributes(mod)
mod$df.residual

# Challenge 7
# fixing errors
# gapminder[gapminder$year = 1957,]
gapminder[gapminder$year == 1957,]
# gapminder[,-1:4]
gapminder[,-c(1:4)]
# gapminder[gapminder$lifeExp > 80]
gapminder[gapminder$lifeExp > 80,]
# gapminder[1, 4, 5]
gapminder[1, c(4, 5)]
# gapminder[gapminder$year == 2002 | 2007,]
gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
gapminder[gapminder$year %in% c(2002, 2007),]

# Challenge 8
# 1. gapminder is a dataframe and needs two columns. gapminder[1:20, ] is a subset with the first 20 rows and all columns
gapminder_small <- gapminder[c(1:9, 19:23),] # gives rows 1 through 9 and 19 through 23
```


## Exercise 3: Creating Publication Quality Graphics

```{r}
library("ggplot2")
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
# scatterplot

ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country, color=continent)) +
  geom_line()
#line chart

ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country, color=continent)) +
  geom_line() + geom_point()
#line chart with points

ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
  geom_line(aes(color=continent)) + geom_point()
# color only applies to line, not points

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = 0.5) + scale_x_log10()
# log transformation

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + scale_x_log10() + geom_smooth(method="lm")
# chart with "best fit" line

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + scale_x_log10() + geom_smooth(method="lm", size=1.5)
# making a thicker line

starts.with <- substr(gapminder$country, start = 1, stop = 1)
az.countries <- gapminder[starts.with %in% c("A", "Z"), ]
ggplot(data = az.countries, aes(x = year, y = lifeExp, color=continent)) +
  geom_line() + facet_wrap( ~ country)
# using faceting for individual country charts

ggplot(data = az.countries, aes(x = year, y = lifeExp, color=continent)) +
  geom_line() + facet_wrap( ~ country) +
  xlab("Year") + ylab("Life expectancy") + ggtitle("Figure 1") +
  scale_colour_discrete(name="Continent") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
# modifying text

# Challenge 1
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()

# Challenge 2
ggplot(data = gapminder, aes(x = year, y = lifeExp, color=continent)) +
  geom_point()

# Challenge 3
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
 geom_point() + geom_line(aes(color=continent)) # point appears under the line

# Challenge 4a
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
 geom_point(size=3, color="red") + scale_x_log10() +
 geom_smooth(method="lm", size=1.5)

# Challenge 4b
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
geom_point(size=3, pch=17) + scale_x_log10() +
geom_smooth(method="lm", size=1.5)

# Challenge 5
ggplot(data = gapminder, aes(x = gdpPercap, fill=continent)) +
 geom_density(alpha=0.6) + facet_wrap( ~ year) + scale_x_log10()
```

